---
title: "Homework 5"
author: "Caitlin Jeffrey"
date: "March 3, 2021"
output: html_document
---

# Homework 5

## Question 1

### 1. Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.

```{r}
x<-1.1; a<-2.2; b<-3.3

#a. 
z <- x^(a^b)
print(z)
```

```{r}
#b. 
z <- (x^a)^b
print(z)
```

```{r}
#c. 
z <- (3*(x^3))+(2*(x^2))+1
print(z)
```


## Question 2

### Using the rep and seq functions, create the following vectors:

```{r}
#a. (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)

c(seq(from=1, to=8, by=1), seq(7,1,-1))
```

```{r}
#b. (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)

my_vec <- c(1,2,3,4,5)
rep(x=my_vec,times=my_vec)
```

```{r}
#c. (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

my_vec <- c(5,4,3,2,1)
other_vec <- c(1,2,3,4,5)
rep(x=my_vec,times=other_vec)
```

## Question 3

### Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates

```{r}
v <- c(runif(1), runif(1))
print(v)

theta <- atan(v[2]/v[1])
r <- sqrt((v[1]^2)+(v[2]^2))

polar_coordinates<-c(theta, r)
print(polar_coordinates)
```


## Question 4

### Suppose that queue <- c("sheep", "fox", "owl", "ant") and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as:

```{r}
#a.the serpent arrives and gets in line

queue <- c("sheep", "fox", "owl", "ant")
a_q <-c(queue, "serpent")
print(a_q)
```

```{r}
#b.the sheep enters the ark

b_q <-a_q[2:5]
print(b_q)
```

```{r}
#c.the donkey arrives and talks his way to the front of the line

c_q <- c(b_q, "donkey")
c_q <- c("donkey", c_q[1:4])
print(c_q)
```


```{r}
#d.the serpent gets impatient and leaves

d_q <- c(c_q[1:4])
print(d_q)
```


```{r}
#e.the owl gets bored and leaves;

e_q <- c(d_q[1:2], d_q[4])
print(e_q)
```

```{r}
#f. the aphid arrives and the ant invites him to cut in line
f_q <- c(e_q[1:2], "aphid", e_q[3])
print(f_q)
```


```{r}
#g. Finally, determine the position of the aphid in the line.
f_q == "aphid"
which(f_q == "aphid")
```

## Question 5

### Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}
int <- seq(from=1, to=100, by=1)

int[!(!int%%2) + (!int%%3) + (!int%%7)]
```












